Epistemic Semantics in Guarded String Models

### 2020-06-17 tex/draft.tex
polish atom description

Kleene Algebra with Tests is an algebraic theory that is defined by
equations and inequalities, which a has model classes including
guarded string models, relational models, finite models, and matrix
models.  This paper focuses on defining a family of concrete guarded
string algebras, the elements of which are sets of guarded strings.
Definitions and notation mostly follow \cite{kozen2001automata}.
Additional syntax and semantics is included to model multi-agent
epistemic semantics.  Guarded strings over a finite alphabet $E$ are
like ordinary strings, but with truth assignments to a set $T$ of
primitive propositions (primitive tests) alternating with the symbols
from $E$. Following \cite{kozen2001automata}, we call these truth
assignments \emph{atoms}; the set of atoms is notated $\mathcal A$.
In the algebra described by Figure \ref{epikfigure}, $E$ is the set of
events $\{a_1,a_0,b_1,b_0\}$, and in the elevator example, $\{u,d\}$.
In the elevator example, $T$ is $\{p,q\}$, and in the coin example it
is $\{h,t\}$.  We write atoms as the boolean formulae that uniquely
specify the truth assignment. In the elevator example, we have atoms
the vectors $\bar q \bar p$, $\bar q p$, $q \bar p$, $q p$.  In the
coin example, we get $h\bar t$ and $\bar h t$, as well as the bogus
atoms $\bar h \bar t$ and $ht$!  To preclude bogosity, the
\texttt{constraint} declaration in Figure~\ref{epikfigure} describes
the set of \emph{feasible atoms}, which we parse into a formula
$\varphi$ (if there are multiple \texttt{constraint declarations},
$\varphi$ is the conjunction of the parsed formulae). For a formula
$\phi$, the set of feasible atoms is $\mathcal A_\phi$. In the coin
example, the constraint $\varphi$ is $h\bar t + \bar h t$. Since $\bar
h t \models \varphi$ and $h \bar t \models \varphi$, they are
permitted feasible atoms, and since $\bar h \bar t$ and $h t$ do not
satisfy $\phi$ they are not permitted.

@@

\begin{example}   \label{gs1}
\begin{tabular}[t]{ll}
$v_1$ & $\bar p \bar q ~u~ \bar p q ~u~ q \bar p ~d~ \bar p q ~u~ q \bar p$ \\
$v_2$ & $\bar p \bar q ~u~ \bar p q ~u~ p\bar q ~u~ pq$ \\   
$w_1$ & $\bar h t$ \\
$w_2$ & $\bar h t ~a_0~ \bar h t$ \\  
$w_3$ & $\bar h t ~a_0~ \bar h t ~b_0~ \bar h t$ \\
\end{tabular}
\end{example}

@@

\begin{example} \label{fusion1}
\hspace{-2mm}\( \bar p q ~u~ \bar p q ~u~ p \bar q ~d~ \bar p q \cdot \bar p q ~u~ p \bar q = \bar p q ~u~ \bar p q ~u~ p \bar q ~d~ \bar p q ~u~ p \bar q \)

\hspace{-2mm}\( \bar p \bar q ~u~ \bar p q ~u~ p \bar q ~d~ \bar p q \cdot p \bar q ~u~ pq = \mbox{\em{undefined}} \)

\hspace{-2mm}\(\bar h t \cdot \bar h t ~a_0~ \bar h t = \bar h t ~a_0~ \bar h t \)

\hspace{-2mm}\(\bar h t \cdot h \bar t ~a_1~ h \bar t = \mbox{\em{undefined}} \)
\end{example}

@@

\begin{example} \label{events1}
\(
\begin{array}[t]{l|c}
  e & \llbracket e \rrbracket^{\varphi, P} \\ \hline  
  u & \{ \bar p \bar q ~u~ \bar p q, \bar p q ~u ~p \bar q,  p \bar q ~u~ pq\}\\
  d & \{ \bar p q ~d~ \bar p \bar q, p \bar q~d~\bar p q, p q~d~ p \bar q\}\\
  a_1 & \{h \bar t ~a_1~ h \bar t \}\\
  a_0 & \{t \bar h ~a_0~ t \bar h\}\\
  b_1 & \{h\bar t ~b_1~ h \bar t \}\\
  b_0 & \{t \bar h ~b_0~ t \bar h \}\\
\end{array}
\)
\end{example}

>>>>

Kleene Algebra with Tests is an algebraic theory that is defined by
equations and inequalities, which a has model classes including
guarded string models, relational models, finite models, and matrix
models.  This paper focuses on defining a family of concrete guarded
string algebras, the elements of which are sets of guarded strings.
Definitions and notation mostly follow \cite{kozen2001automata}.
Additional syntax and semantics is included to model multi-agent
epistemic semantics.  Guarded strings over a finite alphabet $E$ are
like ordinary strings, but with truth assignments to a set $T$ of
primitive propositions (primitive tests) alternating with the symbols
from $E$. Following \cite{kozen2001automata}, we call these truth
assignments \emph{atoms}; the set of atoms is notated $\mathcal A$.
In the algebra described by Figure \ref{epikfigure}, $E$ is the set of
events $\{a_1,a_0,b_1,b_0\}$, and in the elevator example, $\{u,d\}$.
In the elevator example, $T$ is $\{p,q\}$, and in the coin example it
is $\{h,t\}$.  We write atoms as the boolean formulae that uniquely
specify the truth assignment. In the elevator example, we have atoms
$\bar q \bar p$, $\bar q p$, $q \bar p$, $q p$, for notational
simplicity we will write these as $\hat0$, $\hat1$, $\hat2$, $\hat3$
respectivelyto indicate the floor each atom represents. In the coin
example, we get $h\bar t$ and $\bar h t$ (for which we use the
shorthand $H$ and $T$), as well as the bogus atoms $\bar h \bar t$ and
$ht$!  To preclude bogosity, the \texttt{constraint} declaration in
Figure~\ref{epikfigure} describes the set of \emph{feasible atoms},
which we parse into a formula $\varphi$ (if there are multiple
\texttt{constraint} declarations, $\varphi$ is the conjunction of the
parsed formulae in each). For a formula $\phi$, the set of feasible atoms is
$\mathcal A_\phi$. In the coin example, the constraint $\varphi$ is
$h\bar t + \bar h t$. Since $\bar h t \models \varphi$ and $h \bar t
\models \varphi$, they are permitted feasible atoms, and since $\bar h
\bar t$ and $h t$ do not satisfy $\phi$ they are not permitted.

@@

%% \begin{example}   \label{gs1}
%% \begin{tabular}[t]{ll}
%% $v_1$ & $\bar p \bar q ~u~ \bar p q ~u~ q \bar p ~d~ \bar p q ~u~ q \bar p$ \\
%% $v_2$ & $\bar p \bar q ~u~ \bar p q ~u~ p\bar q ~u~ pq$ \\   
%% $w_1$ & $\bar h t$ \\
%% $w_2$ & $\bar h t ~a_0~ \bar h t$ \\  
%% $w_3$ & $\bar h t ~a_0~ \bar h t ~b_0~ \bar h t$ \\
%% \end{tabular}
%% \end{example}
\begin{example}   \label{gs1}
\begin{tabular}[t]{ll}
$v_1$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 ~u~ \hat 2$ \\
$v_2$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~u~ \hat 3$ \\   
$w_1$ & $H$ \\
$w_2$ & $H ~a_0~ H$ \\  
$w_3$ & $T~a_0~ T ~b_0~ T$ \\
\end{tabular}
\end{example}

@@

\begin{example} \label{fusion1}
\hspace{-2mm}\(\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 \cdot \hat 1 ~u~ \hat 2 = \hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 ~u~ \hat 2 \)

\hspace{-2mm}\( \hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 \cdot \hat 2 ~u~ \hat 3 = \mbox{\em{undefined}} \)

\hspace{-2mm}\(H \cdot H ~a_0~ H = H ~a_0~ H \)

\hspace{-2mm}\(T \cdot T ~a_1~ T = \mbox{\em{undefined}} \)
\end{example}

@@

\begin{example} \label{events1}
\(
\begin{array}[t]{l|c}
  e & \llbracket e \rrbracket^{\varphi, P} \\ \hline  
  u & \{ \hat 0 ~u~ \hat 1, \hat 1 ~u~ \hat 2,  \hat 2 ~u~ \hat 3\}\\
  d & \{ \hat 1 ~d~ \hat 0, \hat 2~d~\hat 1, \hat 3 ~d~ \hat 2\}\\
  a_1 & \{H ~a_1~ H\}\\
  a_0 & \{T ~a_0~ T\}\\
  b_1 & \{H ~b_1~ H \}\\
  b_0 & \{T ~b_0~ T \}
\end{array}
\)
\end{example}
